---
title: "Analyses speech coding over daylong recordings"
output: pdf_document
	toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

## Key research questions

## Basic (not to redo)
getting an idea of difficulty via coding time
for j in ${folder}*/*m1.TextGrid
do
 date=`date -r $j`
 echo $j $date >> ~/Desktop/times.txt
done
```{r get-codtime, eval=FALSE}
reliab_folder="/Users/acristia/Documents/habilis_backup/PROJECTS/1-developing/WACK-all/coded/"
read.table("~/Desktop/times.txt")->x
x$date=as.Date(paste0(x$V4,x$V3,x$V7),format = "%d%b%Y")
x$dt=as.POSIXlt(paste(x$date,x$V5))
x[order(x$dt),]->y
y$dur=as.numeric(as.character(difftime(c(y$dt[-1],NA),y$dt,units="secs")))
hist(y$dur,xlim=c(0,1000),breaks=10000)
y$dur[y$dur>1000]<-NA
y$name=gsub(".*/","",y$V1,perl=T)
y$name=gsub(".TextGrid","",y$name,fixed=T)
write.table(y[c("name","dur","date","dt")],"../derivedFiles/code_dur.txt",row.names=F)
#data.frame(table(x$date))->quant
#quant$order<-row.names(quant)
#cor.test(quant$order,quant$Freq)
#max(quant$order)*7*60/sum(quant$Freq)
#5*7*60/sum(quant$Freq[dim(quant)[1]-4:dim(quant)[1]])
```

## Reliability


```{r create-dataset,eval=FALSE}
dirs=dir(path=reliab_folder)
all=NULL
empty=NULL
for(thisdir in dirs){
	csvs=dir(path=paste0(reliab_folder,thisdir),pattern="csv")
	for(thisf in csvs){
  	read.csv(paste0(reliab_folder,thisdir,"/",thisf))->thiscsv
	 bits=strsplit(thisf,"_")
	 if(dim(thiscsv)[1]>0){
  	  all=rbind(all,cbind(bits[[1]][1],bits[[1]][4],bits[[1]][5],bits[[1]][6],thiscsv))
	 }else{
  	  empty=rbind(empty,cbind(bits[[1]][1],bits[[1]][4],bits[[1]][5],bits[[1]][6]))
	 }
	}	
}

names(all)[1:4]<-c("file","start","end","coder")

#minor clean
all$tier_broad=as.character(all$tier)
all$tier_broad=gsub(" ","",all$tier)
all$tier_broad[grep("parl",all$tier_broad)]<-"2parlou+"
all$tier_broad[grep("oi",all$tier_broad)]<-"Loi-fai-2e"
all$tier_broad[all$tier_broad %in% c("C1","C2")]<-"OCH"
all$tier_broad[all$tier_broad %in% c("FA1","FA2")]<-"FA"
all$tier_broad[all$tier_broad %in% c("MA1","MA2")]<-"MA"

all$age=NA
all$age[all$tier_broad %in% c("CHI","OCH")]<-"C"
all$age[all$tier_broad %in% c("FA","MA")]<-"A"

all$parole=NA
all$parole[all$tier_broad %in% c("OCH","FA","MA")]<-"otherSpeech"


all$coder<-gsub(".csv","",all$coder)
all$dur=all$tmax-all$tmin

write.table(all, "../derivedFiles/final.txt",row.names=F)
```

## Comparing old & new

Clarify old and new

```{r moveNew,eval=F}
dir(path="../../nam-j/_1minute/",pattern=".wav")->old
dir(path="../../nam-j/new1min/",pattern=".wav")->new
added=new[new %in% old]

for(thisf in added){
  file.copy(paste0("../../nam-j/new1min/",thisf),paste0("../../nam-j/added/",thisf))
  thisf=gsub(".wav",".csv",thisf)
  file.copy(paste0("../../nam-j/new1min/",thisf),paste0("../../nam-j/added/",thisf))
   thisf=gsub(".csv",".TextGrid",thisf)
  file.copy(paste0("../../nam-j/new1min/",thisf),paste0("../../nam-j/added/",thisf))

}
```

## Reading in current results

New version of tg, compose

uses code from https://stackoverflow.com/questions/41839268/sum-overlapping-non-overlapping-time-intervals-in-r by Jonathan von Schroeder

```{r doAll,eval=F}

reliab_folder="../../nam-j/all/"
csvs=dir(path=reliab_folder,pattern="csv")
all=NULL
for(thisf in csvs){
  #csvs[1]->thisf
  read.csv(paste0(reliab_folder,thisf))->thiscsv
  bits=strsplit(thisf,"_")
  if(dim(thiscsv)[1]==0) thiscsv[1,]<-NA
  all=rbind(all,cbind(thisf, bits[[1]][1],bits[[1]][2],bits[[1]][3],thiscsv))
}

names(all)<-c("File","child","date","chunkstart",	"beg",	"speakerID",	 "type",	"end")

#tier name clean
all$speakerID=gsub("*","",all$speakerID,fixed=T)
all$speakerID=gsub(" ","",all$speakerID,fixed=T)
table(all$speakerID)
all$speakerID[all$speakerID == "Autre"]<-"Noise"
all$speakerID[all$speakerID == "FA2"]<-"FA3"
all$speakerID[all$speakerID == "FA1"]<-"FA2"
all$speakerID[all$speakerID == "MOT"]<-"FA1"
all$speakerID[all$speakerID %in% c("+2parl","2POPMT","2parlou+")]<-"XOL" #overlap bet 2 or more speakers hard to separate
all$speakerID[all$speakerID %in% c("LF2P","Loin","Loin-faible")]<-"SP" #far away speech, "second plane" 
all$speakerID=gsub("^C","UC",all$speakerID,perl=T)
all$speakerID[all$speakerID %in% c("UCHI")]<-"CHI"
table(all$speakerID)

all$type[all$type=="1\177"]<-1
all$type[all$type=="xx"]<-"x"
all$type[all$type=="X"]<-"x"
all$type[all$type %in% c(" ","\177")]<-NA
all$type[all$type %in% c("1&","11")]<-"1"
table(all$type)

all$chunk=as.numeric(as.character(gsub(".csv","",all$chunk,fixed=T)))

all$dur=all$end-all$beg

all$recn=substr(all$date,9,9)

all$date=as.Date(paste(substr(all$date,1,4),substr(all$date,5,6),substr(all$date,7,8),sep="-"))


# #classes
# all$tier_broad=as.character(all$tier)
# all$tier_broad=gsub(" ","",all$tier)
# all$tier_broad[grep("parl",all$tier_broad)]<-"2parlou+"
# all$tier_broad[grep("oi",all$tier_broad)]<-"Loi-fai-2e"
# all$tier_broad[all$tier_broad %in% c("C1","C2")]<-"OCH"
# all$tier_broad[all$tier_broad %in% c("FA1","FA2")]<-"FA"
# all$tier_broad[all$tier_broad %in% c("MA1","MA2")]<-"MA"
# 
# all$age=NA
# all$age[all$tier_broad %in% c("CHI","OCH")]<-"C"
# all$age[all$tier_broad %in% c("FA","MA")]<-"A"
# 
# all$parole=NA
# all$parole[all$tier_broad %in% c("OCH","FA","MA")]<-"otherSpeech"

write.table(all,"../derivedFiles/line_per_segment.txt", row.names=F,quote=T,sep="\t")
```

## do summaries per child & file

```{r doSums,eval=F}
###### local functions
extract_interval_as_vector <- function(df) {
  as.vector(t(subset(df,select=c('beg','end'))))
}

sum_length <- function(v) {
  sum(v[seq(2,length(v),2)]-v[seq(1,length(v),2)])
}

sum_length_of_overlaps <- function(v1,v2) {
  id <- rep(c(1,0),c(length(v1),length(v2)))
  m <- rbind(id,1-id,c(v1,v2))
  m <- m[,order(m[3,])]
  idx <- which(cumsum(m[1,]) %% 2 & cumsum(m[2,]) %% 2)
  if(length(idx)) sum(sapply(idx,function(i) m[3,i+1]-m[3,i]))
  else 0
}

all_df=all[all$speakerID!="Noise" &    all$speakerID!="SP",]

sums=NULL
for(thisf in levels(all$File)){
  subset(all_df,File==thisf)->df
  all_names <- unique(df$speakerID)
  if(length(all_names)>1){
   # print(length(all_names>1))
    combs <- combn(all_names,2)
    for(i in 1:ncol(combs)) {
     df.sub1 <- subset(df,speakerID == combs[1,i])
     df.sub2 <- subset(df,speakerID == combs[2,i])
      int1= extract_interval_as_vector(df.sub1)
      int2= extract_interval_as_vector(df.sub2)
      l1 <- sum_length(int1) #sum(df.sub1$duration)
      l2 <- sum_length(int2) #sum(df.sub2$duration)
      ol=sum_length_of_overlaps(int1,int2)
      x=cbind(df[1,1:4],combs[1,i],l1, T)
      names(x)<-c(names(df)[1:4],"source","dur","clean")
      y=cbind(df[1,1:4],combs[2,i],l2, T)
      names(y)<-c(names(df)[1:4],"source","dur","clean")
      z=cbind(df[1,1:4],paste(combs[1,i],combs[2,i]),ol ,F)
      names(z)<-c(names(df)[1:4],"source","dur","clean")
      sums=rbind(sums, x,y,z)
    } 
  } else{ 
    x=cbind(df[1,1:4],all_names[1],sum(df$dur),T) 
    names(x)<-c(names(df)[1:4],"source","dur","clean")
    sums=rbind(sums, x)}
}

write.table(sums,"../derivedFiles/line_per_chunk.txt", row.names=F,quote=T,sep="\t")
```

## Add age info 

move age adding above to all, so we don't need to do it twice (for all, for sums).

```{r add-age-all}
##PROBLEM!!! FOR SURE XOANAN IS AMBIGUOUS
#KXAO IS TOO (2 KIDS HAVE IT AS LAST NAME)
#GAKU NOT PRESENT
read.table("../derivedFiles/ages.txt",header=T)->ages
cbind(stack(ages[,c(1:3)]),ages$shortname)->x
gsub("X","",x$ind)->x$ind
x[,3]=paste0(x[,2],x[,3])
x=x[,c(1,3)]
names(x)<-c("age","key")

all$key=paste0(substr(gsub("-","",all$date),1,6),all$child)
merge(all,x,all.x=F,all.y=F)->all

write.table(all,"../derivedFiles/line_per_segment_age.txt", row.names=F,quote=T,sep="\t")

```

```{r add-age-sums}


sums$key=paste0(substr(sums$date,1,6),sums$child)
merge(sums,x,all.x=F,all.y=F)->sums

write.table(sums,"../derivedFiles/line_per_chunk_age.txt", row.names=F,quote=T,sep="\t")

```



## Initial analyses: check 1

The goal of these initial analyses is to confirm strong predictions, as a way to validate the analyses.

First, how much do CHI and FAN vocalize as a function of CHI's age?


```{r prod}

plot(sums$dur~sums$age,subset=c(sums$source=="CHI"),xlab="Age",ylab="Total voc time per chunk",main="CHI")

library(gplots)

plotmeans(sums$dur~sums$age,subset=c(sums$source=="CHI"),xlab="Age",ylab="Total voc time per chunk",main="CHI")

aggregate(sums$dur,by=list(sums$source,sums$age,sums$child,sums$clean),mean)->means

names(means)<-c("source","age","child","clean","dur")

plot(means$dur~means$age,subset=c(means$source=="CHI" & means$clean),xlab="Age",ylab="Average total voc time per chunk",main="CHI")

plot(means$dur~means$age,subset=c(means$source=="FA1"& means$clean),xlab="Age",ylab="Average total voc time per chunk",main="FA1")


```
I had expected FA1 to vocalize MORE as children age, yet I observe the opposite. Perhaps that was not the right prediction to make? I also filtered (as the final code above shows) for clean (non-overlaps) only, and for FA1 this pattern is exacerbated.

I also expected CHI to vocalize MORE as they age, yet I observe no trend.  This was a strong prediction that I am not willing to revise, and which instead makes me suspicious of the data. This may indicate that there is a bias to detect CHI regardless of age. One way that this suspicion could be lifted to a certain extent is the proportion of linguistic vocs going up with age. (Not totally, as the coder knew the age so could have been biased to hear more ling segments as a function of age.)

## Initial analyses: check 2

Strong prediction that should be verified: higher proportion of ling as kids age.

```{r prod-ling}
aggregate(all$dur,by=list(all$child,all$age,all$speakerID,all$type),sum,na.rm=T)->sumdur

names(sumdur)<-c("child","age","speakerID","type","totdur")
hist(sumdur$totdur) #total speaking time over a whole day
#TO IMPROVE -- add number of chunks so as to have total length of coding per chunk

sumdur[sumdur$speakerID=="CHI" ,]->sumchi
sumchi$unique=paste0(sumchi$child,sumchi$age)
sumchi[sumchi$type==0,5:6]->chinonling
names(chinonling)[1]<-"nonling"
sumchi[sumchi$type==1,-3]->chiling
names(chiling)[4]<-"ling"
merge(chiling,chinonling,all.x=T,all.y=T)->sumchi2
sumchi2$tot=sumchi2$ling+sumchi2$nonling
sumchi2$prop=sumchi2$ling/sumchi2$tot


plot(sumchi2$prop~sumchi2$age,cex=(sumchi2$tot/max(sumchi2$tot,na.rm=T)))

summary(lm(prop~age+(1/child),data=sumchi2))
summary(lm(prop~age+(1/child),data=sumchi2,subset=c(!(sumchi2$age<10 & sumchi2$prop>.7))))

write.table(sumchi2,"../derivedFiles/propling_age.txt",row.names=F,sep="\t")
```
That is a pretty massive effect with one exception of an 8mo have a really high ling voc rate. Notice that this child has a rather large voc quantity.

So these results are more encouraging, although as mentioned above it could be due to the coder's expectations...

## Check 3: Stability in behavior over days

We often recorded 3 days -- how stable is vocal behavior across days within the same participant (CHI & FA1)?

```{r prod-ling-stab-chi}
aggregate(all$dur,by=list(all$child,all$age,all$speakerID,all$type,all$date),sum,na.rm=T)->sumdur_all

names(sumdur_all)<-c("child","age","speakerID","type","date","totdur")

subset(sumdur_all,speakerID == "CHI" & type==1)->sumdur
sumdur$key=paste0(sumdur$child,sumdur$age,sumdur$speakerID)
sumdur[order(sumdur$key),]->sumdur

sumdur[sumdur$key=="kxa9CHI",]

names(table(sumdur$key))[table(sumdur$key)>=2]->include
sumdur[sumdur$key %in% include,]->sumdur


twodays=NULL
for(eachkid in levels(as.factor(sumdur$key))){
  subset(sumdur,key==eachkid)->x
  xx=NULL
  for(i in 1:2) xx=rbind(xx,cbind(x[1,-6],mean(x[seq(i,dim(x)[1],by=2),"totdur"]),i))
  twodays=rbind(twodays,xx)
}
names(twodays)[7:8]<-c("avgtotdur","day")

plot(twodays[twodays$day==2,"avgtotdur"]~twodays[twodays$day==1,"avgtotdur"],xlab="Odd days",ylab="Even days", main="Avg tot voc duration CHI")
abline(lm(twodays[twodays$day==2,"avgtotdur"]~twodays[twodays$day==1,"avgtotdur"]))

cor.test(twodays[twodays$day==2,"avgtotdur"],twodays[twodays$day==1,"avgtotdur"],method="s")
```

Weak correlation, not super convincing.


```{r prod-ling-stab-fa1}
subset(sumdur_all,speakerID == "FA1" & type==1)->sumdur
sumdur$key=paste0(sumdur$child,sumdur$age,sumdur$speakerID)
sumdur[order(sumdur$key),]->sumdur

names(table(sumdur$key))[table(sumdur$key)>=2]->include
sumdur[sumdur$key %in% include,]->sumdur


twodays=NULL
for(eachkid in levels(as.factor(sumdur$key))){
  subset(sumdur,key==eachkid)->x
  xx=NULL
  for(i in 1:2) xx=rbind(xx,cbind(x[1,-6],mean(x[seq(i,dim(x)[1],by=2),"totdur"]),i))
  twodays=rbind(twodays,xx)
}
names(twodays)[7:8]<-c("avgtotdur","day")

plot(twodays[twodays$day==2,"avgtotdur"]~twodays[twodays$day==1,"avgtotdur"],xlab="Odd days",ylab="Even days", main="Avg tot voc duration CHI")
abline(lm(twodays[twodays$day==2,"avgtotdur"]~twodays[twodays$day==1,"avgtotdur"]))

cor.test(twodays[twodays$day==2,"avgtotdur"],twodays[twodays$day==1,"avgtotdur"],method="s") 
```

Even weaker correlation, not super convincing AT ALL.

### correlation on sums



```{r sums}
#create sums per tier, coder, and chunk
sums=aggregate(all$dur,by=list(all$coder,all$start,all$tier_broad),sum)
names(sums)<-c("coder","start","tier","sumdur")
sums$uid=paste(sums$start,sums$tier)
sums_m1<-sums[sums$coder=="m1",]
sums.y<-sums[sums$coder=="m2",]
joint=merge(sums_m1,sums.y,by="uid")
names(joint)<-gsub(".x","_m1",names(joint))
names(joint)<-gsub(".y",".y",names(joint))


#create sums per parole, coder, and chunk
sums=aggregate(all$dur,by=list(all$coder,all$start,all$parole),sum)
names(sums)<-c("coder","start","tier","sumdur")
sums$uid=paste(sums$start,sums$tier)
sums_m1<-sums[sums$coder=="m1",]
sums.y<-sums[sums$coder=="m2",]
joint_p=merge(sums_m1,sums.y,by="uid")
names(joint_p)<-gsub(".x","_m1",names(joint_p))
names(joint_p)<-gsub(".y",".y",names(joint_p))

#create sums per parole, coder, and chunk
sums=aggregate(all$dur,by=list(all$coder,all$start,all$age),sum)
names(sums)<-c("coder","start","tier","sumdur")
sums$uid=paste(sums$start,sums$tier)
sums_m1<-sums[sums$coder=="m1",]
sums.y<-sums[sums$coder=="m2",]
joint_a=merge(sums_m1,sums.y,by="uid")
names(joint_a)<-gsub(".x","_m1",names(joint_a))
names(joint_a)<-gsub(".y",".y",names(joint_a))

pdf("comparison2.pdf")

plot(joint$sumdur_m1~joint$sumdur.y)
for(thislevel in levels(as.factor(joint$tier_m1))){
  plot(joint$sumdur_m1 [joint$tier_m1 == thislevel] ~joint$sumdur.y [joint$tier_m1 == thislevel], main=thislevel)
  if(length(joint$sumdur_m1 [joint$tier_m1 == thislevel]) > 3) text(mean(joint$sumdur.y [joint$tier_m1 == thislevel]),mean(joint$sumdur_m1 [joint$tier_m1 == thislevel]),round(cor.test(joint$sumdur.y [joint$tier_m1 == thislevel], joint$sumdur_m1 [joint$tier_m1 == thislevel])$estimate,3))
}


for(thislevel in levels(as.factor(joint_p$tier_m1))){
  plot(joint_p$sumdur_m1 [joint_p$tier_m1 == thislevel] ~joint_p$sumdur.y [joint_p$tier_m1 == thislevel], main=thislevel)
  if(length(joint_p$sumdur_m1 [joint_p$tier_m1 == thislevel]) > 3)   text(mean(joint_p$sumdur.y [joint_p$tier_m1 == thislevel]),mean(joint_p$sumdur_m1 [joint_p$tier_m1 == thislevel]),round(cor.test(joint_p$sumdur.y [joint_p$tier_m1 == thislevel], joint_p$sumdur_m1 [joint_p$tier_m1 == thislevel])$estimate,3))
}


for(thislevel in levels(as.factor(joint_a$tier_m1))){
  plot(joint_a$sumdur_m1 [joint_a$tier_m1 == thislevel] ~joint_a$sumdur.y [joint_a$tier_m1 == thislevel], main=thislevel)
  if(length(joint_a$sumdur_m1 [joint_a$tier_m1 == thislevel]) > 3)  text(mean(joint_a$sumdur.y [joint_a$tier_m1 == thislevel]),mean(joint_a$sumdur_m1 [joint_a$tier_m1 == thislevel]),round(cor.test(joint_a$sumdur.y [joint_a$tier_m1 == thislevel], joint_a$sumdur_m1 [joint_a$tier_m1 == thislevel])$estimate,3))
}

dev.off()
```

# repeated
## Basic (not to redo)
getting an idea of difficulty via coding time
for j in ${folder}*/*m1.TextGrid
do
 date=`date -r $j`
 echo $j $date >> ~/Desktop/times.txt
done
```{r get-codtime, eval=FALSE}
reliab_folder="/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/2017_reliab/coded/"
read.table("~/Desktop/times.txt")->x
x$date=as.Date(paste0(x$V4,x$V3,x$V7),format = "%d%b%Y")
x$dt=as.POSIXlt(paste(x$date,x$V5))
x[order(x$dt),]->y
y$dur=as.numeric(as.character(difftime(c(y$dt[-1],NA),y$dt,units="secs")))
hist(y$dur,xlim=c(0,1000),breaks=10000)
y$dur[y$dur>1000]<-NA
y$name=gsub(".*/","",y$V1,perl=T)
y$name=gsub(".TextGrid","",y$name,fixed=T)
write.table(y[c("name","dur","date","dt")],"../derivedFiles/code_dur.txt",row.names=F)
#data.frame(table(x$date))->quant
#quant$order<-row.names(quant)
#cor.test(quant$order,quant$Freq)
#max(quant$order)*7*60/sum(quant$Freq)
#5*7*60/sum(quant$Freq[dim(quant)[1]-4:dim(quant)[1]])
```

## Reliability


```{r create-dataset,eval=FALSE}
dirs=dir(path=reliab_folder)
all=NULL
empty=NULL
for(thisdir in dirs){
	csvs=dir(path=paste0(reliab_folder,thisdir),pattern="csv")
	for(thisf in csvs){
  	read.csv(paste0(reliab_folder,thisdir,"/",thisf))->thiscsv
	 bits=strsplit(thisf,"_")
	 if(dim(thiscsv)[1]>0){
  	  all=rbind(all,cbind(bits[[1]][1],bits[[1]][4],bits[[1]][5],bits[[1]][6],thiscsv))
	 }else{
  	  empty=rbind(empty,cbind(bits[[1]][1],bits[[1]][4],bits[[1]][5],bits[[1]][6]))
	 }
	}	
}

names(all)[1:4]<-c("file","start","end","coder")

#minor clean
all$tier_broad=as.character(all$tier)
all$tier_broad=gsub(" ","",all$tier)
all$tier_broad[grep("parl",all$tier_broad)]<-"2parlou+"
all$tier_broad[grep("oi",all$tier_broad)]<-"Loi-fai-2e"
all$tier_broad[all$tier_broad %in% c("C1","C2")]<-"OCH"
all$tier_broad[all$tier_broad %in% c("FA1","FA2")]<-"FA"
all$tier_broad[all$tier_broad %in% c("MA1","MA2")]<-"MA"

all$age=NA
all$age[all$tier_broad %in% c("CHI","OCH")]<-"C"
all$age[all$tier_broad %in% c("FA","MA")]<-"A"

all$parole=NA
all$parole[all$tier_broad %in% c("OCH","FA","MA")]<-"otherSpeech"


all$coder<-gsub(".csv","",all$coder)
all$dur=all$tmax-all$tmin

write.table(all, "../derivedFiles/final.txt",row.names=F)
```

## Comparing old & new

Clarify old and new

```{r moveNew,eval=F}
dir(path="../../nam-j/_1minute/",pattern=".wav")->old
dir(path="../../nam-j/new1min/",pattern=".wav")->new
added=new[new %in% old]

for(thisf in added){
  file.copy(paste0("../../nam-j/new1min/",thisf),paste0("../../nam-j/added/",thisf))
  thisf=gsub(".wav",".csv",thisf)
  file.copy(paste0("../../nam-j/new1min/",thisf),paste0("../../nam-j/added/",thisf))
   thisf=gsub(".csv",".TextGrid",thisf)
  file.copy(paste0("../../nam-j/new1min/",thisf),paste0("../../nam-j/added/",thisf))

}
```
