---
title: "Analyses speech patterns in daylong recordings across multiple sites"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
redoALL=FALSE ## LOOK HERE!!!

```
# Instructions for use

If this is first time that you are running this script, change the redoALL parameter to TRUE so that the first set of chunks, under generate main databases, is recalculated. Make sure you've created a folder called "derivedFiles" inside dalohumacosp, since the databases will be stored in that folder. (The folder is not pushed to github.)


# Generate main databases 

## Prep

Convert textgrids into csv's.

```{bash convert,eval=redoALL, include=redoALL}
final_folder="/Users/acristia/Documents/namvan_coded/" #where coded files lie
/Applications/Praat.app/Contents/MacOS/Praat --run "1_compareCoding_praat.PraatScript" $final_folder
mkdir ../derivedFiles/coded
mv $final_folder/*/*.csv ../derivedFiles/coded/
```

## Reading in current coding

Read in all csv's and clean up.

```{r doAll,eval=redoALL, include=redoALL}
csv_folder="../derivedFiles/coded/"
csvs=dir(path=csv_folder,pattern="csv")
all=NULL
for(thisf in csvs){
  #csvs[1]->thisf
  read.csv(paste0(csv_folder,thisf))->thiscsv
  bits=strsplit(thisf,"_")
  if(length(bits[[1]])==5){child=bits[[1]][1]; date=bits[[1]][2]; recn=bits[[1]][3]; time=bits[[1]][4]} else {child=bits[[1]][1]; date=bits[[1]][2]; recn=1; time=bits[[1]][3]}
  if(dim(thiscsv)[1]==0) thiscsv[1,]<-NA
  all=rbind(all,cbind(thisf, child,date,recn, time,thiscsv))
}
#getting error invalid factor level, NA generated -- not sure why

names(all)<-c("File","child","date","recn","chunkstart",	"beg",	"speakerID",	 "type",	"end")

#tier name clean
all$speakerID=gsub("*","",all$speakerID,fixed=T)
all$speakerID=gsub(" ","",all$speakerID,fixed=T)
table(all$speakerID)
table(all$child,all$speakerID)
all$speakerID[all$speakerID == "Autre"]<-"Noise"
all$speakerID[all$speakerID == "FA2"]<-"FA3"
all$speakerID[all$speakerID == "FA1"]<-"FA2"
all$speakerID[all$speakerID == "MOT"]<-"FA1"
all$speakerID[all$speakerID %in% c("+2parl","2POPMT","2parlou+")]<-"XOL" #overlap bet 2 or more speakers hard to separate
all$speakerID[all$speakerID %in% c("LF2P","Loin","Loin-faible")]<-"SP" #far away speech, "second plane" 
all$speakerID=gsub("^C","UC",all$speakerID,perl=T)
all$speakerID[all$speakerID %in% c("UCHI")]<-"CHI"
table(all$speakerID)

all$type[all$type=="1\177"]<-1
all$type[all$type=="xx"]<-"x"
all$type[all$type=="X"]<-"x"
all$type[all$type %in% c(" ","\177")]<-NA
all$type[all$type %in% c("1&","11")]<-"1"
table(all$type)

all$end[all$end>240 & !is.na(all$end)]<-240

all$dur=all$end-all$beg

all$date=as.Date(paste(substr(all$date,1,4),substr(all$date,5,6),substr(all$date,7,8),sep="-"))

all$site=ifelse(substr(all$File,1,1)=="v","Van","Nam")

write.table(all,"../derivedFiles/line_per_segment.txt", row.names=F,quote=T,sep="\t")

read.table("../derivedFiles/info.txt",header=T)->ages
cbind(stack(ages[,c(1:3)]),ages$Code)->x
gsub("X","",x$ind)->x$ind
x[,3]=paste0(x[,2],x[,3])
x=x[,c(1,3)]
names(x)<-c("age","key")

all$key=paste0(substr(gsub("-","",all$date),1,6),all$child)
all$key[all$site=="Van"]<-paste0("201703",all$child[all$site=="Van"])
merge(all,x,all.x=T,all.y=F)->all

all$chiage=paste0(all$child,all$age)

all$chidate=paste0(all$child,all$date)

all$totNchunks=NA
for(eachchidate in levels(all$chidate)) all$totNchunks[all$chidate==eachchidate]<-length(levels(factor(all$File[all$chidate==eachchidate])))

write.table(all,"../derivedFiles/line_per_segment_age.txt", row.names=F,quote=T,sep="\t")

```

## Draw summaries per child & file

The goal is to find out how much clean and overlapping speech there is by all of the speakers, summing within each file. This chunk uses code from https://stackoverflow.com/questions/41839268/sum-overlapping-non-overlapping-time-intervals-in-r by Jonathan von Schroeder.

```{r doSums,eval=redoALL, include=redoALL}
###### local functions
extract_interval_as_vector <- function(df) {
  as.vector(t(subset(df,select=c('beg','end'))))
}

sum_length <- function(v) {
  sum(v[seq(2,length(v),2)]-v[seq(1,length(v),2)])
}

sum_length_of_overlaps <- function(v1,v2) {
  id <- rep(c(1,0),c(length(v1),length(v2)))
  m <- rbind(id,1-id,c(v1,v2))
  m <- m[,order(m[3,])]
  idx <- which(cumsum(m[1,]) %% 2 & cumsum(m[2,]) %% 2)
  if(length(idx)) sum(sapply(idx,function(i) m[3,i+1]-m[3,i]))
  else 0
}

all_df=all[all$speakerID!="Noise" &    all$speakerID!="SP" & !is.na(all$beg),]

sums=NULL
for(thisf in levels(all$File)){
  subset(all_df,File==thisf)->df
  all_names <- unique(df$speakerID)
  if(length(all_names)>1){
   # print(length(all_names>1))
    combs <- combn(all_names,2)
    for(i in 1:ncol(combs)) {
     df.sub1 <- subset(df,speakerID == combs[1,i])
     df.sub2 <- subset(df,speakerID == combs[2,i])
      int1= extract_interval_as_vector(df.sub1)
      int2= extract_interval_as_vector(df.sub2)
      l1 <- sum_length(int1) #sum(df.sub1$duration)
      l2 <- sum_length(int2) #sum(df.sub2$duration)
      ol=sum_length_of_overlaps(int1,int2)
      x=cbind(df[1,1:4],combs[1,i],l1, T)
      names(x)<-c(names(df)[1:4],"source","dur","clean")
      y=cbind(df[1,1:4],combs[2,i],l2, T)
      names(y)<-c(names(df)[1:4],"source","dur","clean")
      z=cbind(df[1,1:4],paste(combs[1,i],combs[2,i]),ol ,F)
      names(z)<-c(names(df)[1:4],"source","dur","clean")
      sums=rbind(sums, x,y,z)
    } 
  } else{ 
    x=cbind(df[1,1:4],all_names[1],sum(df$dur),T) 
    names(x)<-c(names(df)[1:4],"source","dur","clean")
    sums=rbind(sums, x)}
}

sums[!is.na(sums$child),]->sums
#we remove all the NA files CHECK

sums$key=paste0(substr(gsub("-","",sums$date),1,6),sums$child)
sums$key[substr(sums$child,1,1)=="v"]<-paste0("201703",sums$child[substr(sums$child,1,1)=="v"])

read.table("../derivedFiles/info.txt",header=T)->ages
cbind(stack(ages[,c(1:3)]),ages$Code)->x
gsub("X","",x$ind)->x$ind
x[,3]=paste0(x[,2],x[,3])
x=x[,c(1,3)]
names(x)<-c("age","key")

merge(sums,x,all.x=T,all.y=F)->sums

sums$totNchunks=NA
for(eachchivisit in levels(factor(sums$key))) sums$totNchunks[sums$key==eachchivisit]<-length(levels(factor(sums$File[sums$key==eachchivisit])))

sums$maxtot=sums$totNchunks*60

sums$propTot=sums$dur/60

write.table(sums,"../derivedFiles/line_per_chunk.txt", row.names=F,quote=T,sep="\t")
```

## Generating Proportions & quantities of CHI vocalizations of different types

```{r dopropCHI,eval= redoALL, include=redoALL}
aggregate(all$dur,by=list(all$child,all$age,all$speakerID,all$type,all$date),sum,na.rm=T)->sumdur
names(sumdur)<-c("child","age","speakerID","type","date","totdur")
sumdur[sumdur$speakerID=="CHI" ,]->sumchi

sumchi$chidate=paste0(sumchi$child,sumchi$date)
sumchi[sumchi$type==0,c("chidate","totdur")]->chinonling
colnames(chinonling)[2]<-"nonling"
sumchi[sumchi$type==1,]->chiling
colnames(chiling)[dim(chiling)[2]-1]<-"ling"
merge(chiling,chinonling,all.x=T,all.y=T)->sumchi
sumchi$ling[is.na(sumchi$ling)]<-0
sumchi$nonling[is.na(sumchi$ling)]<-0
sumchi$tot=sumchi$ling+sumchi$nonling
sumchi$prop=sumchi$ling/sumchi$tot
sumchi$chag=paste0(sumchi$child,sumchi$age)

sumchi$nchunks=NA
for(eachchidate in levels(factor(all$chidate))) sumchi$nchunks[sumchi$chidate==eachchidate]<- length(levels(factor(all$File[all$chidate==eachchidate])))

sumchi$ling.CFRL=sumchi$ling/(sumchi$nchunks*60)
sumchi$nonling.CFRL=sumchi$nonling/(sumchi$nchunks*60)
sumchi$all.CFRL=sumchi$ling.CFRL+sumchi$nonling.CFRL
  
write.table(sumchi,"../derivedFiles/propling_age.txt",row.names=F,sep="\t")
```

## Generating Proportion & quantities of FA1 linguistic vocalizations

```{r dopropsFA1,eval= redoALL, include=redoALL}
aggregate(all$dur,by=list(all$child,all$age,all$speakerID,all$type,all$date),sum,na.rm=T)->sumdur
names(sumdur)<-c("child","age","speakerID","type","date","totdur")
sumdur[sumdur$speakerID=="FA1",]->summot

summot$chidate=paste0(summot$child,summot$date)
summot[summot$type==0,c("chidate","totdur")]->motnonling
colnames(motnonling)[2]<-"nonling"
summot[summot$type==1,]->motling
colnames(motling)[dim(motling)[2]-1]<-"ling"
merge(motling,motnonling,all.x=T,all.y=T)->summot
summot$ling[is.na(summot$ling)]<-0
summot$nonling[is.na(summot$ling)]<-0
summot$tot=summot$ling+summot$nonling
summot$prop=summot$ling/summot$tot
summot$chag=paste0(summot$child,summot$age)

summot$nchunks=NA
for(eachchidate in levels(factor(all$chidate))) summot$nchunks[summot$chidate==eachchidate]<- length(levels(factor(all$File[all$chidate==eachchidate])))

summot$ling.CFRL=summot$ling/(summot$nchunks*60)
summot$nonling.CFRL=summot$nonling/(summot$nchunks*60)
summot$all.CFRL=summot$ling.CFRL+summot$nonling.CFRL


write.table(summot,"../derivedFiles/propling_age_MOT.txt",row.names=F,sep="\t")
```



# Methodological analyses

## How much data are there?

We read in the version of the file that has one line per coded segment. That is, if the child vocalized 3 times in a given minute, an adult 2 times, and another child once, then there will be 6 segments for that *minute*. This is the most fine-grained format of representation, and thus too fine-grained for some of the quantification questions we have. So we will derive tables as needed.

First, we look at how many coded minutes there are for each child and date.

```{r qfy}
read.table("../derivedFiles/line_per_segment_age.txt",header=T)->all
levels(all$child)-> kids
levels(all$chidate)-> days

#number of coded minutes per child
data.frame(table(all$chidate,all$File))->x
x[x$Freq!=0,]->x

#number of coded days per child
data.frame(table(all$child,all$chidate))->y
y[y$Freq!=0,]->y

```

There are `r length(kids)` in all, including `r length(kids[substr(kids,1,1)=="v"])` from Vanuatu and the remaining from Namibia. There are a total of `r length(days)` coded, of which `r length(days[substr(days,1,1)=="v"])` from Vanuatu. Although the two sites have roughly the same number of children (coded -- there are more uncoded children in Vanuatu), more days have been recorded in Namibia:

```{r} 
table(y$Var1)
```

The two sites and groups of kids vary in recording length, which is equivalent to number of minutes coded since there is 1 minute per hour that has been coded. This is clear in the following histograms of number of minutes coded separated by site.

```{r mincoded}
hist(x$Freq[substr(x$Var1,1,1)!="v"],main="Namibia: Total nb of minutes coded")

hist(x$Freq[substr(x$Var1,1,1)=="v"],main="Vanuatu: Total nb of minutes coded")
```

The following table shows the number of minutes coded by recording day.

`r table(x$Var1)`

I am suspicious of the low number of coded minutes for van4 and va10. It is also surprising to see some very high numbers, like that of uebn2017-03-09... We should check those recordings to see whether something went wrong in the concatenation. 



# Theory-relevant analyses
```{r prodsfunctions}
dokidageplot<-function(sumchi,dvname,fancydvname){
  #dvname="prop";fancydvname="prop"
  #dvname="ling.CFRL";fancydvname=""

  plot(sumchi[,dvname]~sumchi$age,type="n",main="",yaxt="n",ylab="",xaxt="n",xlab="")
  mtext(fancydvname,2,line=1)
  axis(2,line=-1,tick=F)
  mtext("Age (months)",1,line=1)
  axis(1,line=-1,tick=F)
  
  mycols=rainbow(length(levels(factor(sumchi$chag))))
names(mycols)<-levels(factor(sumchi$chag))

for(kidag in levels(factor(sumchi$chag))){ 
  age=mean(sumchi$age[sumchi$chag==kidag],na.rm=T)
  prop=mean(sumchi[sumchi$chag==kidag,dvname],na.rm=T)
  err=sd(sumchi[sumchi$chag==kidag,dvname],na.rm=T)
  text(age,prop,substr(kidag,1,3),col=mycols[kidag])
  lines(c(age,age),c(prop-err,prop+err),col="gray")
}
 
  for(kid in levels(factor(sumchi$child))){ 
  props = aggregate(sumchi[sumchi$child==kid,dvname],by=list(sumchi$age[sumchi$child==kid]),mean,na.rm=T)
  for(thisline in 1:(dim(props)[1]-1)) lines(c(props$Group.1[thisline],props$Group.1[thisline+1]),c(props$x[thisline],props$x[thisline+1]),lty=2,col="gray")
}
}

dokidageplot_points<-function(sumchi,dvname,fancydvname){
    #dvname="prop";fancydvname="prop"
mycols=rainbow(length(levels(factor(sumchi$child))))
names(mycols)<-levels(factor(sumchi$child))


  
    plot(sumchi[,dvname]~sumchi$age,cex=(sumchi[,dvname]/max(sumchi[,dvname],na.rm=T)),main="Data per recording day",xlab="Age (months)",ylab=dvname)
text(sumchi$age,sumchi[,dvname],as.character(sumchi$child),cex=.5,pos=4,col=ifelse(substr(kid,1,1)=="v","black",mycols))

}
```

In all of the following analyses, we will consider four possible dependent measures, all of which control for the fact that we have a lot more coded data for some kids than others:

    - The quantity of linguistic vocalizations (controlling for recording length)
    - The quantity of non-linguistic vocalizations (controlling for recording length)
    - The quantity of vocalizations (summing linguistic and non-linguistic) (controlling for recording length)
    - The proportion of vocalizations that are linguistic, out of all vocalizations

When we say controlling for recording length we mean that these are proportions of the time that has been coded where CHI produces ling vocs. For instance, if kid produced 200 secs of ling vocalizations in a given day, of which 1000 secs have been coded, then they will show up below as .2. We don't need to do that for the last DV, since the length of vocalizations is already in the denominator.

## How much does CHI vocalize as a function of age?

Focus on linguistic vocalizations:

```{r prod-age-ling}
read.table("../derivedFiles/propling_age.txt",header=T)->sumchi
subset(sumchi,substr(sumchi$child,1,1)!="v")->sumchi
pdf("ling_age.pdf", height=10, width=15,pointsize = 36)
  par(mar=c(2.5, 2.5, 1, 1))
dokidageplot(sumchi,"ling.CFRL","CHI voc quantity")
dev.off()
```

Now non-linguistic vocalizations:

```{r prod-age-nonling}
dokidageplot(sumchi,"nonling.CFRL","Quantity of CHI non-ling vocs (controlling for rec length)")
```

Now total vocalizations:

```{r prod-age-all}
dokidageplot(sumchi,"all.CFRL","Quantity of CHI vocs (controlling for rec length)")
```

And finally, the ratio of ling to all vocs:

```{r prod-age-prop}
dokidageplot(sumchi,"prop","Proportion of CHI  vocs that are linguistic")
```

## How much does MOTHER vocalize as a function of age?

Please note that we believe our "mother" coding to be a lot less reliable in Vanuatu than Namibia. For adults, it doesn't make a lot of sense to look at proportion of vocs that are linguistic, so we just plot quantity of linguistic and quantity total (ling + nonling).

Focus on linguistic vocalizations:

```{r prod-age-ling-mot}

read.table("../derivedFiles/propling_age_MOT.txt",header=T)->summot


dokidageplot(summot,"ling.CFRL","Quantity of MOT ling vocs (controlling for rec length)")
```

Now total vocalizations:

```{r prod-age-all-mot}
dokidageplot(summot,"all.CFRL","Quantity of MOT vocs (controlling for rec length)")
```

## How much input is there, as a function of age?

Since overheard versus child-directed has not yet been coded, this is total input

```{r input-age,eval=F}
read.table("../derivedFiles/line_per_chunk.txt",header=T)->sums

aggregate(sums$dur,by=list(sums$key,sums$source,sums$maxtot,sums$age),sum,na.rm=T)->sum_visit
colnames(sum_visit)<-c("key","source","maxtot","age","tot")
sum_visit$tot.CFRL=sum_visit$tot/sum_visit$maxtot

all_sources=as.character(levels(sum_visit$source))
nonCHI=all_sources[-grep("CHI",all_sources)]

nonOL=all_sources[-grep(" ",all_sources)]
nonOL=nonOL[-grep("XOL",nonOL)]
nonOLnonCHI=nonOL[-grep("CHI",nonOL)]



aggregate(sum_visit$tot.CFRL[sum_visit$source %in% nonOLnonCHI],by=list(sum_visit$key[sum_visit$source %in% nonOLnonCHI],sum_visit$age[sum_visit$source %in% nonOLnonCHI]),sum)->tot_talking_nonOL
colnames(tot_talking_nonOL)<-c("key","age","tot")

aggregate(sum_visit$tot.CFRL[sum_visit$source %in% nonCHI],by=list(sum_visit$key[sum_visit$source %in% nonCHI],sum_visit$age[sum_visit$source %in% nonCHI]),sum)->tot_talking
colnames(tot_talking)<-c("key","age","tot")


plot(tot_talking$tot ~ tot_talking$age,col=ifelse(substr(as.character(tot_talking$key),7,7)=="v","red","blue"),main="Total non-CHI talking with overlap")

plot(tot_talking_nonOL$tot ~ tot_talking_nonOL$age,col=ifelse(substr(as.character(tot_talking$key),7,7)=="v","red","blue"),main="Total non-CHI talking, NO overlap")
```
## How much of the input comes from mother versus other adults versus other kids?

```{r props-sources,eval=F}
aggregate(sum_visit$tot.CFRL[sum_visit$source %in% nonOLnonCHI],by=list(sum_visit$key[sum_visit$source %in% nonOLnonCHI],sum_visit$source[sum_visit$source %in% nonOLnonCHI],sum_visit$age[sum_visit$source %in% nonOLnonCHI]),sum)->tot_talking_nonOL
colnames(tot_talking_nonOL)<-c("key","source","age","tot")

aggregate(sum_visit$tot.CFRL[sum_visit$source %in% nonCHI],by=list(sum_visit$key[sum_visit$source %in% nonCHI],sum_visit$source[sum_visit$source %in% nonCHI],sum_visit$age[sum_visit$source %in% nonCHI]),sum)->tot_talking
colnames(tot_talking)<-c("key","source","age","tot")
```

## Is there a relationship between input quantity and output quantity?

Considering all input

Considering only non-OL input

Considering only MOT



# TO DO:
- identify what I could possibly have meant by: "CHI. voc auqntm speech auqntm ratio"
- add recording onset time
- only then describe quant over hours

# Stability section

```{r prod-ling-stab-chi}
read.table("../derivedFiles/propling_age.txt", header=T)->sumchi
selvars=c("ling","nonling","tot","prop","ling.CFRL","nonling.CFRL")

twodays=NULL
for(eachkid in names(table(sumchi$chag))[table(sumchi$chag)>1 & substring(names(table(sumchi$chag)),1,1)!="v"]){
  subset(sumchi,chag==eachkid)->thischidat
  odd_days=cbind(thischidat[1,c("child","age","chag")],
             t( colMeans(thischidat[seq(1,dim(thischidat)[1],by=2),selvars],na.rm=T)),
              sum(thischidat[seq(1,dim(thischidat)[1],by=2),c("nchunks")],na.rm=T),1)
  names(odd_days)[10:11]<-c("nchunks","day")
  
   even_days=cbind(thischidat[1,c("child","age","chag")],
             t( colMeans(thischidat[seq(2,dim(thischidat)[1],by=2),selvars],na.rm=T)),
              sum(thischidat[seq(2,dim(thischidat)[1],by=2),c("nchunks")],na.rm=T),2)
  names(even_days)[10:11]<-c("nchunks","day")
    twodays=rbind(twodays, odd_days,even_days ) 
}

#thisdv="ling.CFRL"
for(thisdv in "ling.CFRL"){
  temp=table(twodays$chag[!is.na(twodays[,thisdv])])
  goodkids=names(temp)[temp==2]
 
   pdf(paste0(thisdv,".pdf"),height=15,width=15,pointsize = 36)
  par(mar=c(2.5, 2.5, 1, 1))

  plot(twodays[twodays$day==2 & twodays$chag %in% goodkids,thisdv]~twodays[twodays$day==1 & twodays$chag %in% goodkids,thisdv],xaxt="n",yaxt="n", main="",xlim=range(twodays[,thisdv],na.rm=T),ylim=range(twodays[,thisdv],na.rm=T), type="n")
  axis(1,tick=F,line=-1)
  axis(2,tick=F,line=-1)
  mtext("Odd days",1,line=1)
  mtext("Even days",2,line=1)

  text(twodays[twodays$day==1 & twodays$chag %in% goodkids,thisdv],twodays[twodays$day==2 & twodays$chag %in% goodkids,thisdv], twodays[twodays$day==2 & twodays$chag %in% goodkids,"chag"],
       col=rainbow(length(twodays$day==2 & twodays$chag %in% goodkids)))
  mylm=lm(twodays[twodays$day==2,thisdv]~twodays[twodays$day==1,thisdv] + (1/twodays$child[twodays$day==2]),subset=c(twodays$chag %in% goodkids))
abline(mylm)
dev.off()

print(summary(mylm))
print(cor.test(twodays[twodays$day==2 & twodays$chag %in% goodkids ,thisdv],twodays[twodays$day==1 & twodays$chag %in% goodkids,thisdv],method="s"))
difvec=abs(twodays[twodays$day==2 & twodays$chag %in% goodkids,thisdv] - twodays[twodays$day==1 & twodays$chag %in% goodkids,thisdv])
print(summary(difvec*100))
print(cbind(as.character(twodays[twodays$day==2 & twodays$chag %in% goodkids,"chag"]),difvec))
} 

```
