---
title: "Deciding how to sample from extracts to avoid empty chunks (with no speech)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compose all the datasets from various sources


```{r readintxt,eval=F}
dir(pattern="*[0-9]")->mydirs
all=NULL
for(thisdir in mydirs) all=rbind(all, cbind(read.table(paste0(thisdir,"/stats_mean_diff.txt")) ) )
all$excl=ifelse(all$V2>350,1,0)
write.table(all,"/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_mean_diff.txt", row.names=F,quote=T,sep="\t")

```

```{r readintestcsv, eval=F}
#argh I've destroyed my local version of these files -- fortunately, saved a backup below
test=NULL
for(thisdir in mydirs) test=rbind(test, cbind(read.csv(paste0(thisdir,"/test.csv")) ) )
#repeats when the random sampler overlaps with human sampling
#test[test$sound_file=="aiku_20160714_1_12800.wav",]
#test$babytalk[duplicated(test$sound_file)]
#remove those dups
test=test[!duplicated(test$sound_file),]
test$babyvoc=NA
test$babyvoc[!is.na(test$babytalk)]=ifelse(test$babytalk>0,"bb","0")[!is.na(test$babytalk)]

merge(all, test,all.x=T,all.y=T,by.x="V1",by.y="sound_file")->x
write.table(x, "/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_info.txt",row.names=T,quote=T,sep=";")
```

```{bash gencsv, eval=F}
/Applicatip/Contents/MacOS/Praat --run "/Users/alejandrinacristia/gitrepos/dalohumacosp/compAna/0docsv.PraatScript" "/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/universal_materials/"
```

```{r readcsvs,eval=F}
dir(path="/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/nam_sampled_human/", pattern="[0-9].csv")->csvs
allcsvs=NULL
for(thisf in csvs) {
  #csvs[1]->thisf
  read.csv(paste0("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/nam_sampled_human/",thisf))->thiscsv
  thiscsv=thiscsv[grep("CHI",thiscsv$tier),]
  if(dim(thiscsv)[1]>0){
  bits=strsplit(thisf,"_")
  if(length(bits[[1]])==3) thisline=c(bits[[1]][1],bits[[1]][2],1,bits[[1]][3]) else thisline=c(bits[[1]][1],bits[[1]][2],bits[[1]][3],bits[[1]][4])
   allcsvs=rbind(allcsvs,cbind(thisf, t(thisline),thiscsv))
    
  }
  }

names(allcsvs)<-c("File","child","date","rec","chunkstart",	"beg",	"speakerID",	 "type",	"end")

#tier name clean
allcsvs$speakerID=gsub("*","",allcsvs$speakerID,fixed=T)
allcsvs$speakerID=gsub(" ","",allcsvs$speakerID,fixed=T)
table(allcsvs$speakerID)

table(allcsvs$type)
allcsvs$type=gsub("\177","",allcsvs$type,fixed=T)
allcsvs$type=gsub(" ","",allcsvs$type,fixed=T)
allcsvs$type[allcsvs$type=="11"]<-1
allcsvs$type[allcsvs$type==""]<-NA
allcsvs$type[allcsvs$type=="x"]<-NA
table(allcsvs$type)

allcsvs$chunkstart=as.numeric(as.character(gsub(".csv","",allcsvs$chunkstart,fixed=T)))

allcsvs$dur=allcsvs$end-allcsvs$beg

allcsvs$date=as.Date(paste(substr(allcsvs$date,1,4),substr(allcsvs$date,5,6),substr(allcsvs$date,7,8),sep="-"))

write.table(allcsvs,"/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_human_line_per_segment.txt", row.names=F,quote=T,sep="\t")

aggregate(allcsvs$dur,by=list(allcsvs$File,allcsvs$type),sum)->trsums

trsums[trsums$Group.2==0,c(1,3)]->nl
colnames(nl)<-c("file","nl")
trsums[trsums$Group.2==1,c(1,3)]->ling
colnames(ling)<-c("file","l")
merge(nl,ling,all.x=T,all.y=T)->trsumty

write.table(trsumty,"/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_human_line_per_file.txt", row.names=F,quote=T,sep="\t")

```

```{bash ldc,eval=F}
cd /Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/nam_sampled_ldc_SAD
for j in */*.lab; do
  nspeech=`cat $j | grep " speech" | wc -l`
  if [[ $nspeech -gt 0 ]]
  then
    echo $j 1 >> ldc_speech.txt
  else
    echo $j 0 >> ldc_speech.txt
  fi
done

```
```{r read_ldc,eval=F}
read.table("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/nam_sampled_ldc_SAD/ldc_speech.txt")->ldc
ldc$V1=gsub(".*/","",ldc$V1)
ldc$V1=gsub("lab","",ldc$V1)
colnames(ldc)<-c("file","sp")
write.table(ldc,"/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_ldc_all.txt", row.names=F,quote=T,sep="\t")
```

```{r addhuman,eval=F}
read.table("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_human_line_per_file.txt",header=T)->human
read.table("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_info.txt",header=T,sep=";")->means
colnames(means)[1:2]<-c("file","meandif")
gsub("wav","",means$file)->means$file
gsub("csv","",human$file)->human$file

merge(means,human,all.x=T,all.y=T)->humm

read.table("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_ldc.txt",header=T)->ldc

merge(humm,ldc,all.x=T,all.y=T)->humml
humml[humml$file!="child_voice.",]->humml
write.table(humml,"/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_humml.txt", row.names=F,quote=T,sep="\t")

```


## check results using LDC

```{r checkLDC}
read.table("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_humml.txt",header=T)->x
x$sp[!is.na(x$sp)]<-"speech_found"
x$sp[is.na(x$sp)]<-"no"
x$there.is.ling<-ifelse(x$l>0,"yes","no_ling")
x$there.is.ling[is.na(x$l) & !is.na(x$babyvoc)]<-"no_ling"
x$there.is.nonling<-ifelse(x$nl>0,"yes","no_non_ling")
x$there.is.nonling[is.na(x$nl) & !is.na(x$babyvoc)]<-"no_non_ling"

# CO OCCURRENCE WITH ANY TYPE OF VOC
table(x$sp,x$babyvoc)->mytab
print(mytab)
  precision_ldc=mytab[2,2]/sum(mytab[2,])
  recall_ldc=mytab[2,2]/sum(mytab[,2])
print(precision_ldc)
print(recall_ldc)

# CO OCCURRENCE WITH LING VOC
table(x$sp,x$there.is.ling)->mytab
print(mytab)
  precision=mytab[2,2]/sum(mytab[2,])
  recall=mytab[2,2]/sum(mytab[,2])
print(precision)
print(recall)

# CO OCCURRENCE WITH NONLING VOC
table(x$sp,x$there.is.nonling)->mytab
print(mytab)
  precision=mytab[2,2]/sum(mytab[2,])
  recall=mytab[2,2]/sum(mytab[,2])
print(precision)
print(recall)


```


## check results using loudness


```{r checkLoudnessPM}
# CO OCCURRENCE WITH ANY TYPE OF VOC
table(x$excl,x$babyvoc)->mytab
print(mytab)
  precision=mytab[2,2]/sum(mytab[2,])
  recall=mytab[2,2]/sum(mytab[,2])
print(precision)
print(recall)

# CO OCCURRENCE WITH LING VOC
table(x$excl,x$there.is.ling)->mytab
print(mytab)
  precision=mytab[2,2]/sum(mytab[2,])
  recall=mytab[2,2]/sum(mytab[,2])
print(precision)
print(recall)

# CO OCCURRENCE WITH NONLING VOC
table(x$excl,x$there.is.nonling)->mytab
print(mytab)
  precision=mytab[2,2]/sum(mytab[2,])
  recall=mytab[2,2]/sum(mytab[,2])
print(precision)
print(recall)

curves=NULL
for(pm in seq(50,600,10)){
  incl=ifelse(x$meandif>pm,"incl","excl")[!is.na(x$babyvoc)]
  table(incl,x$babyvoc[!is.na(x$babyvoc)])->mytab
  precision=mytab[2,2]/sum(mytab[2,])
  recall=mytab[2,2]/sum(mytab[,2])
  curves=rbind(curves,cbind(pm,precision,recall,sum(mytab[2,])))
}


```

##  Plots all results together

You can also embed plots, for example:

```{r pressure}
plot(curves[,"recall"],curves[,"precision"],type="n",xlim=c(0,1),ylim=c(0,1))
text(curves[,"recall"],curves[,"precision"],curves[,"pm"],cex=.5)
points(recall_ldc,precision_ldc,pch="L",col="red")

plot(curves[,"recall"],curves[,4],type="n", xlab="Precision (gray/orange) /recall (red/black)",ylab="Proportion of chunks included",xlim=c(0,1),ylim=range(c(curves[,4],sum(x$sp=="speech_found")))/sum(mytab))
text(curves[,"recall"],curves[,4]/sum(mytab),curves[,"pm"],cex=.5)
text(curves[,"precision"],curves[,4]/sum(mytab),curves[,"pm"],cex=.5,col="gray")
points(recall_ldc,sum(x$sp=="speech_found")/sum(mytab),pch="L",col="red")
points(precision_ldc,sum(x$sp=="speech_found")/sum(mytab),pch="L",col="orange")
```

##  Prepare selections for actual experiment

Each participant should get a 40 minute experiment

    - 45 mins total
    - 5 mins exit chat
    - 5 mins instruction reading
    - *35 mins for the task*
    - 10s chunk + 5 secs decision --> each trial takes 15 secs
    - in 35 mins, we fit 140 trials
    - max coded is 4*nhours ~ 48 trials with feedback
    - --> we could fit 100 trials (assuming no silence)
    - --> extract 200 trials
    - --> assuming 12h/rec, about 15 trials/h --> one every 4 mins

```{r}
read.table("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/backup_humml.txt",header=T)->testinfo

testinfo[,c("file","babytalk","meandif","nl","l","sp")]->testinfo
testinfo$file<-gsub(".","",testinfo$file,fixed=T)
testinfo$visit=substr(testinfo$file,1,11)
table(testinfo$visit,is.na(testinfo$babytalk))
colnames(testinfo)[1]<-"sound_file"

for(thisvisit in testinfo$visit){ 
  sample(testinfo[testinfo$visit==thisvisit & !is.na(testinfo$babytalk),"sound_file"],40)->withfb
  sample(testinfo[testinfo$visit==thisvisit & !is.na(testinfo$babytalk),"sound_file"],110)->new
  subset(testinfo,sound_file %in% withfb[1:35] | sound_file %in% new)->thiskidtest
  subset(testinfo,sound_file %in% withfb[36:40] )->thiskidtrain
  write.table(thiskidtest,paste0("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/universal_materials_1128/",thisvisit,"/test_final.csv"),row.names=F,sep=",")
  write.table(thiskidtrain,paste0("/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/universal_materials_1128/",thisvisit,"/train_final.csv"),row.names=F,sep=",")
}

```

```{bash mv-files, eval=F}
source="/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/universal_materials_1128/"
target="/Users/alejandrinacristia/habilis_backup/PROJECTS/1-developing/WACK/listeningExperiment/universal/__pool__/"
for thisfolder in ${source}aiku_201607/; do
   cp ${thisfolder}/child_voice.wav $target
   cp ${thisfolder}/*.csv $target
  while read p; do
    thisfile=`echo $p | cut -d, -f1 | tr -d '"'`
     cp ${thisfolder}/${thisfile}.wav $target
  done < ${thisfolder}/train_final.csv
    while read p; do
    thisfile=`echo $p | cut -d, -f1 | tr -d '"'`
     cp ${thisfolder}/${thisfile}.wav $target
  done < ${thisfolder}/test_final.csv
done
```
